---
title: "Efficient, Consistent and Flexible Credit Default Simulation: TRNG & RcppParallel"
author:
  - Riccardo Porreca
  - Roland Schmid
output:
  html_document:
    includes:
      in_header: includes/header.html
      after_body: includes/footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```


### Packages and sources
```{r codebase, results="hide", message=FALSE}
library(dplyr) # portfolio manipulation
# install.packages("dplyr")
library(microbenchmark) # benchmarking
# install.packages("microbenchmark")

if (!require(rTRNG)) {
  devtools::install_github("miraisolutions/rTRNG",
                           dependencies = TRUE)
  # install.packages("devtools")
  library(rTRNG)
}

Rcpp::sourceCpp("code/simulationKernel.cpp") # rTRNG/RcppParallel C++ code
source("code/simulationKernel.R") # R wrapper
source("code/ES99.R") # utility for Expected Shortfall 99% calculation
```


### Data load
```{r data}
# load pf, r, Z
loaded <- load("data/inputDataSmall.RData")
#loaded <- load("data/inputDataBig.RData")
loaded
stopifnot(all(c("pf", "r", "Z") %in% loaded))
J <- nrow(pf)
M <- nrow(Z)
stopifnot(ncol(Z) == J)
stopifnot(all(dim(Z) == dim(r)))
colnames(pf)
dim(Z) %>% setNames(c("M", "J"))
# number of idiosyncratic returns to be simulated for each market scenario
K <- 100
# random seed
s <- 12358
# timer for the net simulation time in simulationKernel
timer <- function(...) {
  system.time(...)["elapsed"]
}
```


### Full simulation by rating
```{r full}
L_rtng <- simulationKernel(pf, Z, r, J, K, agg = pf$rtng, seed = s,
                           timer = timer)
ES99_rtng <- ES99(L_rtng)
sort(ES99_rtng, decreasing = TRUE)
```

### Consistent simulation of the BBB sub-portfolio
```{r BBB}
L_BBB <- simulationKernel(pf %>% filter(rtng == "BBB"), Z, r, J, K, seed = s,
                          timer = timer)
all.equal(c(L_BBB), L_rtng[, "BBB"], check.attributes = FALSE)
ES99_BBB <- ES99(L_BBB)
ES99_BBB
```

### Contribution of individual counterparties to the BBB total ES99
```{r BBBcontr}
pfBBB <- pf %>% filter(rtng == "BBB")
L_jBBBtail <- simulationKernel(pfBBB, Z, r, J, K, 
                               agg = pfBBB$j, mk = tail99(L_BBB), seed = s,
                               timer = timer)
ContrES99_jBBB <- colMeans(L_jBBBtail)
head(ContrES99_jBBB)
all.equal(sum(ContrES99_jBBB), ES99_BBB, check.attributes = FALSE)
```

### Full distribution for the top 3 BBB counterparties (by highest contribution)
```{r BBBtop3}
top3jBBB <- names(sort(ContrES99_jBBB, decreasing = TRUE))[1:3]
pftop3BBB <- pfBBB %>% filter(j %in% top3jBBB) 
L_top3BBB <- simulationKernel(pftop3BBB, Z, r, J, K, 
                              agg = pftop3BBB$j, seed = s,
                              timer = timer)
ES99_top3BBB <- ES99(L_top3BBB)
pftop3BBB %>% 
  select(j, V0, R) %>%
  mutate(ES99 = ES99_top3BBB, ContrES99 = ContrES99_jBBB[top3jBBB],
         Div = ContrES99/ES99, "Contr/V0" = ContrES99/V0)
```

### What-if scenario: top 3 BBB counterparties downgraded (higher PD)
```{r BBBtop3wiL}
pfBBBwi <- pf %>% filter(rtng == "BBB") %>%
  mutate(PD = replace(PD, j %in% top3jBBB, 0.01))
pftop3BBBwi <- pfBBBwi %>% filter(j %in% top3jBBB)
L_top3BBBwi <- simulationKernel(pftop3BBBwi, Z, r, J, K, 
                                agg = pftop3BBBwi$j, seed = s,
                                timer = timer)
```

### What-if scenario: impact on the BBB ES99
```{r BBBtop3wiES}
L_BBBwi <- L_BBB + (rowSums(L_top3BBBwi) - rowSums(L_top3BBB))
ES99_BBBwi <- ES99(L_BBBwi)
cbind(ES99_BBBwi, ES99_BBB)
```

### What-if scenario: new contributions
```{r BBBtop3wiContr}
L_jBBBtailwi <- simulationKernel(pfBBBwi, Z, r, J, K, 
                                 agg = pfBBB$j, mk = tail99(L_BBBwi), seed = s,
                                 timer = timer)
ContrES99_jBBBwi <- colMeans(L_jBBBtailwi)
```


### Benchmarks
```{r benchSetup, results="hide", message=FALSE}
# reduce to a reasonable size for benchmarking
M <- pmin(1000, M)
J <- pmin(600, J)
K <- pmin(10, K)
Z <- Z[1:M, 1:J]
r <- r[1:M, 1:J]
pf <- pf[1:J, ]
# number of repetitions
nBench <- 10

```

#### Number of threads
```{r benchThreads}
withThreadNr <- function(numThreads) {
  RcppParallel::setThreadOptions(numThreads = numThreads)
  invisible(simulationKernel(pf, Z, r, J, K, seed = s))
}
mb <- microbenchmark(withThreadNr(1),
                     withThreadNr(2),
                     withThreadNr(4),
                     times = nBench)
boxplot(mb, unit = "s")
```

#### Size of the sub-portfolio
```{r benchJ}
set.seed(85321)
subJ <- ceiling(J/2^(0:4))
# sub-portfolios with constant PD (in order not to bias the benchmark)
subpfs <-
  lapply(as.list(subJ), 
         function(jlen, pf) {
           pf %>% sample_n(jlen)
         },
         pf %>% mutate(PD = mean(PD))) %>% setNames(subJ)
RcppParallel::setThreadOptions(numThreads = 1)
withSubJ <- function(subJ) {
  invisible(simulationKernel(subpfs[[as.character(ceiling(subJ))]], 
                             Z, r, J, K, seed = s))
}
mb <- microbenchmark(withSubJ(J/2^0),
                     withSubJ(J/2^1),
                     withSubJ(J/2^2),
                     withSubJ(J/2^3),
                     withSubJ(J/2^4),
                     times = nBench)
boxplot(mb, unit = "s")
```

#### Number of sub-simulations
```{r benchMK}
set.seed(85321)
mklen <- ceiling(M*K/2^(0:4))
mks <- 
  lapply(mklen,function(s) {
    sort(sample.int(M*K, s))
  }) %>% setNames(mklen)
RcppParallel::setThreadOptions(numThreads = 1)
withSubMK <- function(mkl) {
  invisible(simulationKernel(pf, Z, r, J, K, seed = s,
                             mk = mks[[as.character(ceiling(mkl))]]))
}
mb <- microbenchmark(withSubMK(M*K/2^0),
                     withSubMK(M*K/2^1),
                     withSubMK(M*K/2^2),
                     withSubMK(M*K/2^3),
                     withSubMK(M*K/2^4),
                     times = nBench)
boxplot(mb, unit = "s")
```


### Source code

#### simulationKernel.R
```{r echo=FALSE}
knitr::read_chunk("code/simulationKernel.R", from = 1, labels = "simKernelR")
```
```{r simKernelR, eval=FALSE}
```

#### ES99.R
```{r echo=FALSE}
knitr::read_chunk("code/ES99.R", from = 1, labels = "ES99")
```
```{r ES99, eval=FALSE}
```

#### simulationKernel.cpp
```{r echo=FALSE}
knitr::read_chunk("code/simulationKernel.cpp", from = 1, labels = "simKernelCpp")
```
```{Rcpp simKernelCpp, eval=FALSE}
```

